---
title: "NJVotersonVote16"
author: "Isabelle Jensen"
date: "2025-06-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(dplyr)
library(readr)
library(tidyr)
library(data.table)
library(knitr)
library(lubridate)
library(janitor)
library(gmodels)
library(gt)
library(kableExtra)
library(ggplot2)
library(haven)

library(survey)
library(dplyr)
library(tidyr)
library(openxlsx) 

# Read the .sav file
NJV <- read_sav("12152NJ16.sav")
#"C:\Users\izzoe\Documents\GitHub\CDCE\NJVoterswork\12152NJ16.sav"



```
Can you do the following (all using the weight)?

-run the frequencies for all of the main variables and create a toplines document that lists the questions and the responses in terms of count and percent (formatted in professional tables)
--Note that the characteristics of 16- & 17-year-olds variables are set up in a weird way and will require some recoding

-run crosstabs of the main variables about the voting age (school board, local, state, national) and key demographics, putting the demogs in the rows and voting age questions in the columns and get the row percentages (formatted in professional tables)

-summarize the key results from the toplines and crabs using a few bullet points


```{r}

View(NJV)

summary(NJV$qd3)

design <- svydesign(ids = ~1, data = NJV, weights = ~weight)


main_vars <- c("q1", "q2", "q3","q4", "q5","q6","q7","q8","q9","q10")  # replace with your actual variable names

# Function to compute weighted freq for one variable
get_weighted_freq <- function(var) {
  formula <- as.formula(paste0("~", var))
  tab <- svytable(formula, design)
  df <- as.data.frame(tab)
  colnames(df) <- c("Response", "Weighted_Count")
  df$Percent <- prop.table(df$Weighted_Count) * 100
  df$Question <- var
  df <- df[, c("Question", "Response", "Weighted_Count", "Percent")]
  return(df)
}

# Apply to all variables
toplines_list <- lapply(main_vars, get_weighted_freq)

# Combine into one data frame
toplines_df <- bind_rows(toplines_list)

toplines_df






# Make sure Percent is rounded and formatted
toplines_df <- toplines_df %>%
  mutate(
    Weighted_Count = round(Weighted_Count, 1),
    Percent = paste0(round(Percent, 1), "\\%")
  )

# Build LaTeX rows
latex_rows <- toplines_df %>%
  mutate(
    QuestionLabel = ifelse(duplicated(Question), " ", Question),
    LaTeX = paste0(
      QuestionLabel, " & ",
      Response, " & ",
      Weighted_Count, " & ",
      Percent, " \\\\"
    )
  ) %>%
  pull(LaTeX)

header <- "\\begin{tabular}{p{6cm}lrr}
\\hline
\\textbf{Question} & \\textbf{Response} & \\textbf{Count} & \\textbf{Percent} \\\\
\\hline"

footer <- "\\hline
\\end{tabular}"

latex_table <- c(header, latex_rows, footer)
latex_table
# Save to .tex file
writeLines(latex_table, "toplines_table.tex")

cat(paste(latex_table, collapse = "\n"))




colnames(NJV)
nrow(NJV)


# Assume your weight variable is called "weight"
library(survey)

# Define a survey design object
njv_design <- svydesign(ids = ~1, weights = ~weight, data = NJV)

# Calculate effective sample size
eff_n <- 1 / mean(weights(njv_design)^2) * sum(weights(njv_design))^2

z <- 1.96
p <- 0.5  # Conservative estimate
moe <- z * sqrt(p * (1 - p) / eff_n)
moe_percent <- round(moe * 100, 1)

moe_percent




summary(NJV$weight)
hist(NJV$weight, main = "Distribution of Weights", xlab = "Weight")


# Raw sample size (unweighted)
n_raw <- nrow(NJV)

# Weighted sample size
n_weighted <- sum(weights(njv_design))

# Effective sample size
eff_n <- 1 / mean(weights(njv_design)^2) * sum(weights(njv_design))^2

# Print all
cat("Raw sample size:", n_raw, "\n")
cat("Weighted sample size:", round(n_weighted), "\n")
cat("Effective sample size:", round(eff_n), "\n")


moe_simple <- 1.96 * sqrt(0.5 * (1 - 0.5) / n_raw)
round(moe_simple * 100, 1)  # Should be ~3.5% for n = 800



library(survey)

# Define the design

NJV$ones <- 1



# Estimate design effect using a dummy proportion variable
njv_design <- svydesign(ids = ~1, weights = ~weight, data = NJV)

deff_obj <- svymean(~ones, njv_design, deff = TRUE)
deff_val <- attr(deff_obj, "deff")[[1]]


n_raw <- nrow(NJV)
eff_n <- n_raw / deff_val

z <- 1.96
p <- 0.5
moe <- z * sqrt(p * (1 - p) / eff_n)
moe_percent <- round(moe * 100, 1)

cat("Design effect:", round(deff_val, 2), "\n")
cat("Effective sample size:", round(eff_n), "\n")
cat("Margin of error: ±", moe_percent, "%\n")


```

## Including Plots

You can also embed plots, for example:

```{r, q11}
library(dplyr)
library(tidyr)
library(survey)

# Step 1: Stack q11_01 to q11_09 into long format
q11_long <- NJV %>%
  select(weight, starts_with("q11_")) %>%
  pivot_longer(cols = starts_with("q11_"), names_to = "slot", values_to = "selected_value") %>%
  filter(!is.na(selected_value))  # keep only valid selections

design_q11 <- svydesign(ids = ~1, data = q11_long, weights = ~weight)

q11_freq <- svytable(~selected_value, design_q11)

# Convert to data frame
q11_df <- as.data.frame(q11_freq)
colnames(q11_df) <- c("Value", "Weighted_Count")

# Calculate percent
q11_df <- q11_df %>%
  mutate(
    Percent = round(100 * Weighted_Count / sum(Weighted_Count), 1)
  )

q11_df

q11_labels <- c(
  "1" = "Mature",
  "2" = "Immature",
  "3" = "Knowledgeable",
  "4" = "Uninformed",
  "5" = "Engaged in the community",
  "6" = "Apathetic",
  "7" = "Easily influenced",
  "8" = "Angry",
  "9" = "Bored"
)

q11_df <- q11_df %>%
  mutate(Label = q11_labels[as.character(Value)]) %>%
  select(Label, Weighted_Count, Percent)

q11_df
```


-run crosstabs of the main variables about the voting age (school board, local, state, national) and key demographics, putting the demogs in the rows and voting age questions in the columns and get the row percentages (formatted in professional tables)
q1 school board
q4 local
q6 state
q8 national

qd2 party rep 1 dem 2 ind 3
qd4 age
qd6 latino 1
qd7 white black asian
qd10 income
qdll man 1 woman 2 else 3

```{r, party crosstab}
#school board party

library(srvyr)
library(dplyr)
library(tidyr)

# 1. Filter qd2 to just 1 and 2
NJV_filtered <- NJV %>%
  filter(qd2 %in% c(1, 2))

# 2. Create survey design with weights
survey_data <- NJV_filtered %>%
  as_survey_design(weights = weight)

# 3. Weighted counts
counts <- survey_data %>%
  group_by(qd2, q1) %>%
  summarize(n = survey_total(na.rm = TRUE), .groups = "drop")

# 4. Weighted row-wise percentages (within qd2)
percents <- survey_data %>%
  group_by(qd2, q1) %>%
  summarize(pct = survey_mean(proportion = TRUE, na.rm = TRUE), .groups = "drop") %>%
  mutate(pct = round(100 * pct, 1))

# 5. Merge counts and percents
combined <- left_join(counts, percents, by = c("qd2", "q1")) %>%
  mutate(cell = paste0(round(n, 1), " (", pct, "%)")) %>%
  select(qd2, q1, cell)

# 6. Pivot to wide format: qd2 as rows, q1 as columns
crosstab_final <- combined %>%
  pivot_wider(names_from = q1, values_from = cell, names_prefix = "q1_")
crosstab_final

library(xtable)
print(xtable(crosstab_final), include.rownames = FALSE, type = "latex")


##all school board

generate_crosstab <- function(df, weight_var, row_var, col_var) {
  df_filtered <- df %>%
    filter(!!rlang::sym(row_var) %in% c(1, 2))  # filter qd2 == 1 or 2

  survey_data <- df_filtered %>%
    as_survey_design(weights = !!rlang::sym(weight_var))

  counts <- survey_data %>%
    group_by(!!rlang::sym(row_var), !!rlang::sym(col_var)) %>%
    summarize(n = survey_total(na.rm = TRUE), .groups = "drop")

  percents <- survey_data %>%
    group_by(!!rlang::sym(row_var), !!rlang::sym(col_var)) %>%
    summarize(pct = survey_mean(proportion = TRUE, na.rm = TRUE), .groups = "drop") %>%
    mutate(pct = round(100 * pct, 1))

  combined <- left_join(counts, percents, by = c(row_var, col_var)) %>%
    mutate(cell = paste0(round(n, 1), " (", pct, "%)")) %>%
    select(!!rlang::sym(row_var), !!rlang::sym(col_var), cell)

  # Pivot to wide
  wide <- combined %>%
    pivot_wider(names_from = !!rlang::sym(col_var), values_from = cell, names_prefix = paste0(col_var, "_"))

  return(wide)
}



ct_q1 <- generate_crosstab(NJV, weight_var = "weight", row_var = "qd2", col_var = "q1")
ct_q4 <- generate_crosstab(NJV, weight_var = "weight", row_var = "qd2", col_var = "q4")
ct_q6 <- generate_crosstab(NJV, weight_var = "weight", row_var = "qd2", col_var = "q6")
ct_q8 <- generate_crosstab(NJV, weight_var = "weight", row_var = "qd2", col_var = "q8")


library(xtable)

print(
  xtable(ct_q1, caption = "Support for School Board Elections by Party"),
  type = "latex",
  include.rownames = FALSE
)

print(
  xtable(ct_q4, caption = "Support for Local Elections by Party"),
  type = "latex",
  include.rownames = FALSE
)

print(
  xtable(ct_q6, caption = "Support for State Elections by Party"),
  type = "latex",
  include.rownames = FALSE
)

print(
  xtable(ct_q8, caption = "Support for National Elections by Party"),
  type = "latex",
  include.rownames = FALSE
)




```
```{r race crosstab}


library(srvyr)
library(dplyr)
library(tidyr)

generate_crosstab <- function(NJV, weight_var, row_var, col_var) {
  
  # Step 1: Filter rows to relevant row variable values (you can customize this before calling)
  NJV_filtered <- NJV
  
  # Step 2: Create survey design
  survey_data <- NJV_filtered %>%
    as_survey_design(weights = !!rlang::sym(weight_var))
  
  # Step 3: Get weighted counts
  counts <- survey_data %>%
    group_by(!!rlang::sym(row_var), !!rlang::sym(col_var)) %>%
    summarize(n = survey_total(na.rm = TRUE), .groups = "drop")
  
  # Step 4: Get weighted percentages (row-wise)
  percents <- survey_data %>%
    group_by(!!rlang::sym(row_var), !!rlang::sym(col_var)) %>%
    summarize(pct = survey_mean(proportion = TRUE, na.rm = TRUE), .groups = "drop") %>%
    mutate(pct = round(100 * pct, 1))
  
  # Step 5: Combine count and percent into one string
  combined <- left_join(counts, percents, by = c(row_var, col_var)) %>%
    mutate(cell = paste0(round(n, 1), " (", pct, "%)")) %>%
    select(!!sym(row_var), !!sym(col_var), cell)
  
  # ✅ Step 6: Ensure all combinations of row/column are included (this is the fix!)
  combined <- combined %>%
    complete(!!sym(row_var), !!sym(col_var), fill = list(cell = "0 (0%)"))
  
  # Step 7: Pivot to wide format for LaTeX table
  wide <- combined %>%
    pivot_wider(names_from = !!sym(col_var), values_from = cell,
                names_prefix = paste0(col_var, "_"))
  
  return(wide)
}

ct_q1_qd7 <- generate_crosstab(
  NJV %>%
    mutate(qd7 = as.numeric(qd7)) %>%      # make sure qd7 is numeric
    filter(qd7 %in% c(1, 2, 3)),           # filter to the 3 values you want
  weight_var = "weight",
  row_var = "qd7",
  col_var = "q1"
)

ct_q1_qd7 <- ct_q1_qd7 %>%
  mutate(qd7 = recode(qd7,
                      `1` = "White",
                      `2` = "Black",
                      `3` = "Asian"))


library(xtable)

print(
  xtable(ct_q1_qd7, caption = "Support for School Board Elections by Race"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)


ct_q4_qd7 <- generate_crosstab(
  NJV %>%
    mutate(qd7 = as.numeric(qd7)) %>%      # make sure qd7 is numeric
    filter(qd7 %in% c(1, 2, 3)),           # filter to the 3 values you want
  weight_var = "weight",
  row_var = "qd7",
  col_var = "q4"
)

ct_q4_qd7 <- ct_q4_qd7 %>%
  mutate(qd7 = recode(qd7,
                      `1` = "White",
                      `2` = "Black",
                      `3` = "Asian"))

print(
  xtable(ct_q4_qd7, caption = "Support for Local Elections by Race"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)



ct_q6_qd7 <- generate_crosstab(
  NJV %>%
    mutate(qd7 = as.numeric(qd7)) %>%      # make sure qd7 is numeric
    filter(qd7 %in% c(1, 2, 3)),           # filter to the 3 values you want
  weight_var = "weight",
  row_var = "qd7",
  col_var = "q6"
)

ct_q6_qd7 <- ct_q6_qd7 %>%
  mutate(qd7 = recode(qd7,
                      `1` = "White",
                      `2` = "Black",
                      `3` = "Asian"))

print(
  xtable(ct_q6_qd7, caption = "Support for State Elections by Race"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)



ct_q8_qd7 <- generate_crosstab(
  NJV %>%
    mutate(qd7 = as.numeric(qd7)) %>%      # make sure qd7 is numeric
    filter(qd7 %in% c(1, 2, 3)),           # filter to the 3 values you want
  weight_var = "weight",
  row_var = "qd7",
  col_var = "q8"
)

ct_q8_qd7 <- ct_q8_qd7 %>%
  mutate(qd7 = recode(qd7,
                      `1` = "White",
                      `2` = "Black",
                      `3` = "Asian"))

print(
  xtable(ct_q8_qd7, caption = "Support for National Elections by Race"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)
```
```{r Latino}
table(NJV$qd6)

ct_q1_qd6 <- generate_crosstab(
  NJV %>%
    mutate(qd6 = as.numeric(qd6)) %>%     # or as_factor if using labels
    filter(qd6 %in% c(1, 2)),          # restrict to valid values
  weight_var = "weight",
  row_var = "qd6",
  col_var = "q1"
)

ct_q1_qd6 <- ct_q1_qd6 %>%
  mutate(qd6 = recode(qd6,
                      `1` = "Latino",
                      `2` = "Non-Latino"))


print(
  xtable(ct_q1_qd6, caption = "Support for School Board Elections by Ethnicity"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)




ct_q4_qd6 <- generate_crosstab(
  NJV %>%
    mutate(qd6 = as.numeric(qd6)) %>%     # or as_factor if using labels
    filter(qd6 %in% c(1, 2)),          # restrict to valid values
  weight_var = "weight",
  row_var = "qd6",
  col_var = "q4"
)

ct_q4_qd6 <- ct_q4_qd6 %>%
  mutate(qd6 = recode(qd6,
                      `1` = "Latino",
                      `2` = "Non-Latino"))


print(
  xtable(ct_q4_qd6, caption = "Support for Local Elections by Ethnicity"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)


ct_q6_qd6 <- generate_crosstab(
  NJV %>%
    mutate(qd6 = as.numeric(qd6)) %>%     # or as_factor if using labels
    filter(qd6 %in% c(1, 2)),          # restrict to valid values
  weight_var = "weight",
  row_var = "qd6",
  col_var = "q6"
)

ct_q6_qd6 <- ct_q6_qd6 %>%
  mutate(qd6 = recode(qd6,
                      `1` = "Latino",
                      `2` = "Non-Latino"))


print(
  xtable(ct_q6_qd6, caption = "Support for State Elections by Ethnicity"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)


ct_q8_qd6 <- generate_crosstab(
  NJV %>%
    mutate(qd6 = as.numeric(qd6)) %>%     # or as_factor if using labels
    filter(qd6 %in% c(1, 2)),          # restrict to valid values
  weight_var = "weight",
  row_var = "qd6",
  col_var = "q8"
)

ct_q8_qd6 <- ct_q8_qd6 %>%
  mutate(qd6 = recode(qd6,
                      `1` = "Latino",
                      `2` = "Non-Latino"))


print(
  xtable(ct_q8_qd6, caption = "Support for National Elections by Ethnicity"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)




ct_q1_qd11 <- generate_crosstab(
  NJV %>%
    mutate(qd11 = as.numeric(qd11)) %>%     # or as_factor if using labels
    filter(qd11 %in% c(1, 2)),          # restrict to valid values
  weight_var = "weight",
  row_var = "qd11",
  col_var = "q1"
)

ct_q1_qd11 <- ct_q1_qd11 %>%
  mutate(qd11 = recode(qd11,
                      `1` = "Men",
                      `2` = "Women"))


print(
  xtable(ct_q1_qd11, caption = "Support for School Board Elections by Gender"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)




ct_q4_qd11 <- generate_crosstab(
  NJV %>%
    mutate(qd11 = as.numeric(qd11)) %>%     # or as_factor if using labels
    filter(qd11 %in% c(1, 2)),          # restrict to valid values
  weight_var = "weight",
  row_var = "qd11",
  col_var = "q4"
)

ct_q4_qd11 <- ct_q4_qd11 %>%
  mutate(qd11 = recode(qd11,
                      `1` = "Men",
                      `2` = "Women"))


print(
  xtable(ct_q4_qd11, caption = "Support for Local Elections by Gender"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)





ct_q6_qd11 <- generate_crosstab(
  NJV %>%
    mutate(qd11 = as.numeric(qd11)) %>%     # or as_factor if using labels
    filter(qd11 %in% c(1, 2)),          # restrict to valid values
  weight_var = "weight",
  row_var = "qd11",
  col_var = "q6"
)

ct_q6_qd11 <- ct_q6_qd11 %>%
  mutate(qd11 = recode(qd11,
                      `1` = "Men",
                      `2` = "Women"))


print(
  xtable(ct_q6_qd11, caption = "Support for State by Gender"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)



ct_q8_qd11 <- generate_crosstab(
  NJV %>%
    mutate(qd11 = as.numeric(qd11)) %>%     # or as_factor if using labels
    filter(qd11 %in% c(1, 2)),          # restrict to valid values
  weight_var = "weight",
  row_var = "qd11",
  col_var = "q8"
)

ct_q8_qd11 <- ct_q8_qd11 %>%
  mutate(qd11 = recode(qd11,
                      `1` = "Men",
                      `2` = "Women"))


print(
  xtable(ct_q8_qd11, caption = "Support for National Election by Gender"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)


```


```{r, income and child}
table(NJV$qd10)#income
table(NJV$qd5)#child or not
summary(NJV$qd4)#age

229/(229+575)



ct_q1_qd10 <- generate_crosstab(
  NJV %>%
    mutate(qd10 = as.numeric(qd10)) %>%     # or as_factor if using labels
    filter(qd10 %in% c(1, 2, 3, 4)),          # restrict to valid values
  weight_var = "weight",
  row_var = "qd10",
  col_var = "q1"
)


print(
  xtable(ct_q1_qd10, caption = "Support for School Board Elections by Income"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)



ct_q4_qd10 <- generate_crosstab(
  NJV %>%
    mutate(qd10 = as.numeric(qd10)) %>%     # or as_factor if using labels
    filter(qd10 %in% c(1, 2, 3, 4)),          # restrict to valid values
  weight_var = "weight",
  row_var = "qd10",
  col_var = "q4"
)


print(
  xtable(ct_q4_qd10, caption = "Support for Local Elections by Income"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)

ct_q6_qd10 <- generate_crosstab(
  NJV %>%
    mutate(qd10 = as.numeric(qd10)) %>%     # or as_factor if using labels
    filter(qd10 %in% c(1, 2, 3, 4)),          # restrict to valid values
  weight_var = "weight",
  row_var = "qd10",
  col_var = "q6"
)


print(
  xtable(ct_q6_qd10, caption = "Support for State Elections by Income"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)


ct_q8_qd10 <- generate_crosstab(
  NJV %>%
    mutate(qd10 = as.numeric(qd10)) %>%     # or as_factor if using labels
    filter(qd10 %in% c(1, 2, 3, 4)),          # restrict to valid values
  weight_var = "weight",
  row_var = "qd10",
  col_var = "q8"
)


print(
  xtable(ct_q8_qd10, caption = "Support for National Elections by Income"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)



ct_q1_qd5 <- generate_crosstab(
  NJV %>%
    mutate(qd5 = as.numeric(qd5)) %>%     # or as_factor if using labels
    filter(qd5 %in% c(1, 2)),          # restrict to valid values
  weight_var = "weight",
  row_var = "qd5",
  col_var = "q1"
)

ct_q1_qd5 <- ct_q1_qd5 %>%
  mutate(qd5 = recode(qd5,
                      `1` = "Has Child",
                      `2` = "Doesn't Have Child"))


print(
  xtable(ct_q1_qd5, caption = "Support for School Board Elections by Parent Status"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)



ct_q4_qd5 <- generate_crosstab(
  NJV %>%
    mutate(qd5 = as.numeric(qd5)) %>%     # or as_factor if using labels
    filter(qd5 %in% c(1, 2)),          # restrict to valid values
  weight_var = "weight",
  row_var = "qd5",
  col_var = "q4"
)

ct_q4_qd5 <- ct_q4_qd5 %>%
  mutate(qd5 = recode(qd5,
                      `1` = "Has Child",
                      `2` = "Doesn't Have Child"))


print(
  xtable(ct_q4_qd5, caption = "Support for Local Elections by Parent Status"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)


ct_q6_qd5 <- generate_crosstab(
  NJV %>%
    mutate(qd5 = as.numeric(qd5)) %>%     # or as_factor if using labels
    filter(qd5 %in% c(1, 2)),          # restrict to valid values
  weight_var = "weight",
  row_var = "qd5",
  col_var = "q6"
)

ct_q6_qd5 <- ct_q6_qd5 %>%
  mutate(qd5 = recode(qd5,
                      `1` = "Has Child",
                      `2` = "Doesn't Have Child"))


print(
  xtable(ct_q6_qd5, caption = "Support for State Elections by Parental Status"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)



ct_q8_qd5 <- generate_crosstab(
  NJV %>%
    mutate(qd5 = as.numeric(qd5)) %>%     # or as_factor if using labels
    filter(qd5 %in% c(1, 2)),          # restrict to valid values
  weight_var = "weight",
  row_var = "qd5",
  col_var = "q8"
)

ct_q8_qd5 <- ct_q8_qd5 %>%
  mutate(qd5 = recode(qd5,
                      `1` = "Has Child",
                      `2` = "Doesn't Have Child"))


print(
  xtable(ct_q8_qd5, caption = "Support for National Elections by Parental Status"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)

```
```{r, age}

summary(NJV$qd4)

NJV <- NJV %>%
  mutate(age_group = case_when(
    qd4 >= 18 & qd4 <= 24 ~ "18–24",
    qd4 >= 25 & qd4 <= 34 ~ "25–34",
    qd4 >= 35 & qd4 <= 44 ~ "35–44",
    qd4 >= 45 & qd4 <= 54 ~ "45–54",
    qd4 >= 55 & qd4 <= 64 ~ "55–64",
    qd4 >= 65 ~ "65+",
    TRUE ~ NA_character_  # for missing or out-of-range values
  ))
table(NJV$age_group)


ct_q1_age <- generate_crosstab(
  NJV %>%
    filter(!is.na(age_group)),  # optional: keep only valid groups
  weight_var = "weight",
  row_var = "age_group",
  col_var = "q1"
)


library(xtable)

print(
  xtable(ct_q1_age, caption = "Support for School Board Elections by Age Group"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)


ct_q4_age <- generate_crosstab(
  NJV %>%
    filter(!is.na(age_group)),  # optional: keep only valid groups
  weight_var = "weight",
  row_var = "age_group",
  col_var = "q4"
)



print(
  xtable(ct_q4_age, caption = "Support for Local Elections by Age Group"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)


ct_q6_age <- generate_crosstab(
  NJV %>%
    filter(!is.na(age_group)),  # optional: keep only valid groups
  weight_var = "weight",
  row_var = "age_group",
  col_var = "q6"
)



print(
  xtable(ct_q6_age, caption = "Support for State Elections by Age Group"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)



ct_q8_age <- generate_crosstab(
  NJV %>%
    filter(!is.na(age_group)),  # optional: keep only valid groups
  weight_var = "weight",
  row_var = "age_group",
  col_var = "q8"
)



print(
  xtable(ct_q8_age, caption = "Support for National Elections by Age Group"),
  include.rownames = FALSE,
  type = "latex",
  caption.placement = "top",
  sanitize.text.function = identity
)

```



```{r education}
colnames(NJV)
#qd3
summary(NJV$qd3)
table(NJV$educ_recode)

# Recode qd3 into new education categories
NJV$educ_recode <- factor(
  case_when(
    NJV$qd3 %in% c(1, 2) ~ "Less than high school",
    NJV$qd3 == 3 ~ "High school graduate",
    NJV$qd3 %in% c(4, 5, 6) ~ "Some college or 2-year degree",
    NJV$qd3 == 7 ~ "4-year college graduate",
    NJV$qd3 == 8 ~ "Graduate degree",
    NJV$qd3 == 9 ~ "Prefer not to say",
    TRUE ~ NA_character_
  ),
  levels = c(
    "Less than high school",
    "High school graduate",
    "Some college or 2-year degree",
    "4-year college graduate",
    "Graduate degree",
    "Prefer not to say"
  )
)

# Crosstab of new education variable
table(NJV$educ_recode, useNA = "ifany")


library(srvyr)
library(dplyr)
library(tidyr)
library(xtable)

# Filter out NA in the education variable
NJV_filtered_educ <- NJV %>%
  filter(!is.na(educ_recode))

# Create survey design
survey_data_educ <- NJV_filtered_educ %>%
  as_survey_design(weights = weight)

# Example for q1
counts_educ <- survey_data_educ %>%
  group_by(educ_recode, q1) %>%
  summarize(n = survey_total(na.rm = TRUE), .groups = "drop")

percents_educ <- survey_data_educ %>%
  group_by(educ_recode, q1) %>%
  summarize(pct = survey_mean(proportion = TRUE, na.rm = TRUE), .groups = "drop") %>%
  mutate(pct = round(100 * pct, 1))

combined_educ <- left_join(counts_educ, percents_educ, by = c("educ_recode", "q1")) %>%
  mutate(cell = paste0(round(n, 1), " (", pct, "%)")) %>%
  select(educ_recode, q1, cell)

crosstab_educ_q1 <- combined_educ %>%
  pivot_wider(names_from = q1, values_from = cell, names_prefix = "q1_")

print(xtable(crosstab_educ_q1, caption = "Support for School Board Elections by Education"), 
      include.rownames = FALSE, type = "latex")



generate_crosstab <- function(df, weight_var, row_var, col_var) {
  df_filtered <- df %>%
    filter(!is.na(!!rlang::sym(row_var)))  # remove NAs

  survey_data <- df_filtered %>%
    as_survey_design(weights = !!rlang::sym(weight_var))

  counts <- survey_data %>%
    group_by(!!rlang::sym(row_var), !!rlang::sym(col_var)) %>%
    summarize(n = survey_total(na.rm = TRUE), .groups = "drop")

  percents <- survey_data %>%
    group_by(!!rlang::sym(row_var), !!rlang::sym(col_var)) %>%
    summarize(pct = survey_mean(proportion = TRUE, na.rm = TRUE), .groups = "drop") %>%
    mutate(pct = round(100 * pct, 1))

  combined <- left_join(counts, percents, by = c(row_var, col_var)) %>%
    mutate(cell = paste0(round(n, 1), " (", pct, "%)")) %>%
    select(!!rlang::sym(row_var), !!rlang::sym(col_var), cell)

  # Pivot to wide
  wide <- combined %>%
    pivot_wider(names_from = !!rlang::sym(col_var), values_from = cell, names_prefix = paste0(col_var, "_"))

  return(wide)
}

ct_q1_educ <- generate_crosstab(NJV, weight_var = "weight", row_var = "educ_recode", col_var = "q1")
ct_q4_educ <- generate_crosstab(NJV, weight_var = "weight", row_var = "educ_recode", col_var = "q4")
ct_q6_educ <- generate_crosstab(NJV, weight_var = "weight", row_var = "educ_recode", col_var = "q6")
ct_q8_educ <- generate_crosstab(NJV, weight_var = "weight", row_var = "educ_recode", col_var = "q8")

# Print to LaTeX
print(xtable(ct_q1_educ, caption = "Support for School Board Elections by Education"), type = "latex", include.rownames = FALSE)
print(xtable(ct_q4_educ, caption = "Support for Local Elections by Education"), type = "latex", include.rownames = FALSE)
print(xtable(ct_q6_educ, caption = "Support for State Elections by Education"), type = "latex", include.rownames = FALSE)
print(xtable(ct_q8_educ, caption = "Support for National Elections by Education"), type = "latex", include.rownames = FALSE)



```


```{r county}

#code 7 is essex
table(NJV$qd8)

NJV$county_group <- factor(
  ifelse(NJV$qd8 == 7, "Essex County", "All Other Counties"),
  levels = c("Essex County", "All Other Counties")
)

ct_q1_county <- generate_crosstab(NJV, weight_var = "weight", row_var = "county_group", col_var = "q1")
ct_q4_county <- generate_crosstab(NJV, weight_var = "weight", row_var = "county_group", col_var = "q4")
ct_q6_county <- generate_crosstab(NJV, weight_var = "weight", row_var = "county_group", col_var = "q6")
ct_q8_county <- generate_crosstab(NJV, weight_var = "weight", row_var = "county_group", col_var = "q8")

print(xtable(ct_q1_county, caption = "Support for School Board Elections by County"), type = "latex", include.rownames = FALSE)
print(xtable(ct_q4_county, caption = "Support for Local Elections by County"), type = "latex", include.rownames = FALSE)
print(xtable(ct_q6_county, caption = "Support for State Elections by County"), type = "latex", include.rownames = FALSE)
print(xtable(ct_q8_county, caption = "Support for National Elections by County"), type = "latex", include.rownames = FALSE)


```





```{r white dems}


colnames(NJV)
#qd2 republican 1 and democrat 2


#qd7, `1` = "White",`2` = "Black",`3` = "Asian"
table(NJV$qd2)
table(NJV$qd7)








# White vs Black Democrats only
NJV$group_white_black_dems <- factor(
  dplyr::case_when(
    NJV$qd2 == 2 & NJV$qd7 == 1 ~ "White Democrats",
    NJV$qd2 == 2 & NJV$qd7 == 2 ~ "Black Democrats",
    TRUE ~ NA_character_  # exclude everyone else
  ),
  levels = c("White Democrats", "Black Democrats")
)

# White Democrats vs White Republicans only
NJV$group_white_dems_reps <- factor(
  dplyr::case_when(
    NJV$qd2 == 2 & NJV$qd7 == 1 ~ "White Democrats",
    NJV$qd2 == 1 & NJV$qd7 == 1 ~ "White Republicans",
    TRUE ~ NA_character_
  ),
  levels = c("White Democrats", "White Republicans")
)

NJV_white_black_dems <- NJV[!is.na(NJV$group_white_black_dems), ]
NJV_white_dems_reps <- NJV[!is.na(NJV$group_white_dems_reps), ]


ct_q1_white_black_dems <- generate_crosstab(NJV_white_black_dems, weight_var = "weight", row_var = "group_white_black_dems", col_var = "q1")
ct_q4_white_black_dems <- generate_crosstab(NJV_white_black_dems, weight_var = "weight", row_var = "group_white_black_dems", col_var = "q4")
ct_q6_white_black_dems <- generate_crosstab(NJV_white_black_dems, weight_var = "weight", row_var = "group_white_black_dems", col_var = "q6")
ct_q8_white_black_dems <- generate_crosstab(NJV_white_black_dems, weight_var = "weight", row_var = "group_white_black_dems", col_var = "q8")

ct_q1_white_dems_reps <- generate_crosstab(NJV_white_dems_reps, weight_var = "weight", row_var = "group_white_dems_reps", col_var = "q1")
ct_q4_white_dems_reps <- generate_crosstab(NJV_white_dems_reps, weight_var = "weight", row_var = "group_white_dems_reps", col_var = "q4")
ct_q6_white_dems_reps <- generate_crosstab(NJV_white_dems_reps, weight_var = "weight", row_var = "group_white_dems_reps", col_var = "q6")
ct_q8_white_dems_reps <- generate_crosstab(NJV_white_dems_reps, weight_var = "weight", row_var = "group_white_dems_reps", col_var = "q8")


print(xtable(ct_q1_white_black_dems, caption = "Support for School Board Elections: White vs. Black Democrats"), type = "latex", include.rownames = FALSE)
print(xtable(ct_q4_white_black_dems, caption = "Support for Local Elections: White vs. Black Democrats"), type = "latex", include.rownames = FALSE)
print(xtable(ct_q6_white_black_dems, caption = "Support for State Elections: White vs. Black Democrats"), 
      type = "latex", include.rownames = FALSE)

print(xtable(ct_q8_white_black_dems, caption = "Support for National Elections: White vs. Black Democrats"), 
      type = "latex", include.rownames = FALSE)







print(xtable(ct_q1_white_dems_reps, caption = "Support for School Board Elections: White Democrats vs. White Republicans"), 
      type = "latex", include.rownames = FALSE)

print(xtable(ct_q4_white_dems_reps, caption = "Support for Local Elections: White Democrats vs. White Republicans"), 
      type = "latex", include.rownames = FALSE)

print(xtable(ct_q6_white_dems_reps, caption = "Support for State Elections: White Democrats vs. White Republicans"), 
      type = "latex", include.rownames = FALSE)

print(xtable(ct_q8_white_dems_reps, caption = "Support for National Elections: White Democrats vs. White Republicans"), 
      type = "latex", include.rownames = FALSE)


table(NJV$qd2)
table(NJV$qd7)

```
