---
title: "NVStadiumVoters"
author: "Isabelle Jensen"
date: "2025-02-11"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
library(readr)
library(tidyr)
library(data.table)
library(knitr)

NVEligibleVoters <- read_csv("VoterList.63884.013125094408/VoterList.ElgbVtr.63884.013125094408.csv")
NVVoteHistory <- read_csv("VoterList.63884.013125094408/VoterList.VtHst.63884.013125094408.csv")
StadiumVoters <- read_excel("AllegiantVotes_2024GeneralElection_VoterIDs (1).xlsx")

colnames(NVEligibleVoters)
glimpse(NVEligibleVoters)
#View(NVEligibleVoters)

colnames(NVVoteHistory)
glimpse(NVVoteHistory)
#View(NVVoteHistory)

colnames(StadiumVoters)

ClarkCountyVoters <- NVEligibleVoters %>%
  filter(`Residential County` == "Clark")

StadiumVoters <- StadiumVoters %>%
  mutate(Allegiant_Voter = "Yes")
#View(StadiumVoters)


ClarkVoterIDs <- ClarkCountyVoters %>%
  select(VoterID)


ClarkCountyVoteHistory <- NVVoteHistory %>%
  semi_join(ClarkVoterIDs, by = "VoterID")
#View(ClarkCountyVoteHistory)


#write.csv(ClarkCountyVoteHistory, "ClarkCountyVoteHistory.csv", row.names = FALSE)
```



The second step is to drop anyone not in Clark County. Note that if any of the IDs don't match you will want to go back and see if they show up elsewhere (might have moved since the election).

-The third step is to match the records of the attached and create a variable for voted at Allegiant Stadium.

-Fourth, do some basic comparisons by age, vote history (especially first time voters), and party reg to compare stadium voters, in-person non-stadium voters, voters by mail, and non-voters. 



```{r combine datasets}
ClarkCountyVoteHistory %>%
  count(VoterID) %>%
  count(n)  # Counts how many voters have 1, 2, 3... elections BEFORE distinct()



#ClarkCountyVoteHistory <- read_csv("ClarkCountyVoteHistory.csv")
#colnames(ClarkCountyVoteHistory)

ClarkCountyVoteHistory <- ClarkCountyVoteHistory %>%
  rename(Election_Date = `Election Date`, Vote_Code = `Vote Code`)
colnames(ClarkCountyVoteHistory)


ClarkCountyVotersandHistory <- ClarkCountyVoters %>%
 left_join(ClarkCountyVoteHistory, by = "VoterID")

View(ClarkCountyVotersandHistory)
colnames(ClarkCountyVotersandHistory)



ClarkCountyVoterwithAllegiant <- ClarkCountyVotersandHistory %>%
  full_join(
    StadiumVoters %>% select(ID_Number, Allegiant_Voter) %>% rename(VoterID = ID_Number),
    by = "VoterID"
  ) %>%
  mutate(Allegiant_Voter = replace_na(Allegiant_Voter, "No"))  # Ensures non-stadium voters are marked "No"

colnames(ClarkCountyVoterwithAllegiant)



fwrite(ClarkCountyVoterwithAllegiant, "ClarkCountyVoterwithAllegiant.csv", row.names = FALSE)
View(ClarkCountyVoterwithAllegiant)


ClarkCountyVoterwithAllegiant <- read_csv("ClarkCountyVoterwithAllegiant.csv")


summary(ClarkCountyVoterwithAllegiant$`Birth Date`)
glimpse(ClarkCountyVoterwithAllegiant$`Birth Date`)

```
clark county voting amounts
1	300256			
2	176476			
3	139314			
4	110424			
5	99003			
6	82866			
7	72829			
8	61572			
9	47326			
10	36775	

```{r analysis of joined data}

ClarkCountyVoterwithAllegiant <- read_csv("ClarkCountyVoterwithAllegiant.csv")

library(lubridate)
library(dplyr)
library(janitor)
ClarkCountyVoterwithAllegiantnames <- ClarkCountyVoterwithAllegiant %>%
  clean_names()
colnames(ClarkCountyVoterwithAllegiantnames)
View(ClarkCountyVoterwithAllegiantnames)

str(ClarkCountyVoterwithAllegiantnames$birth_date)




ClarkCountyVoterwithAllegiantnames <- ClarkCountyVoterwithAllegiantnames %>%
  mutate(
    birth_date = mdy(birth_date)  # Convert from character to Date format
  )

ClarkCountyVoterwithAllegiantnames <- ClarkCountyVoterwithAllegiantnames %>%
  mutate(
    age = as.integer(interval(birth_date, today()) / years(1))  # Compute Age
  )

colnames(ClarkCountyVoterwithAllegiantnames)

ClarkCountyVoterwithAllegiantnames_unique <- ClarkCountyVoterwithAllegiantnames_unique %>%
  group_by(voter_id) %>%
  summarise(age = first(age),  # Keep only one age per voter
            allegiant_voter = first(allegiant_voter)) %>%  # Keep first Yes/No status
  ungroup()

ClarkCountyVoterwithAllegiantnames_unique %>%
  filter(allegiant_voter == "Yes" & !is.na(age)) %>%  # Keep only stadium voters with age
  mutate(age_group = cut(age, breaks = c(18, 30, 45, 60, 100), labels = c("18-29", "30-44", "45-59", "60+"))) %>%
  tabyl(age_group) %>%
  adorn_totals() %>%
  adorn_pct_formatting()



library(gmodels)

CrossTable(ClarkCountyVoterwithAllegiantnames_unique$allegiant_voter, ClarkCountyVoterwithAllegiantnames_unique$age, prop.chisq = FALSE, prop.t = FALSE, prop.r = TRUE)

library(gmodels)
library(dplyr)

# Create age groups
ClarkCountyVoterwithAllegiantnames_unique <- ClarkCountyVoterwithAllegiantnames_unique %>%
  mutate(age_group = cut(age, breaks = c(18, 29, 39, 49, 59, 69, 79, 100),
                         labels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70-79", "80+"),
                         include.lowest = TRUE))

# Crosstab with age groups
CrossTable(ClarkCountyVoterwithAllegiantnames_unique$allegiant_voter, 
           ClarkCountyVoterwithAllegiantnames_unique$age_group, 
           prop.chisq = FALSE, prop.t = FALSE, prop.r = TRUE)

ftable(table(ClarkCountyVoterwithAllegiantnames_unique$allegiant_voter, 
             ClarkCountyVoterwithAllegiantnames_unique$age_group))

crosstabageallegiant <- table(ClarkCountyVoterwithAllegiantnames_unique$allegiant_voter, 
                  ClarkCountyVoterwithAllegiantnames_unique$age_group)

crosstabageallegiant_percent_row <- prop.table(crosstabageallegiant, margin = 1) * 100
print(round(crosstabageallegiant_percent_row, 1))


```
     18-29 30-39 40-49 50-59 60-69 70-79  80+
  No   20.4  19.1  16.2  15.1  14.2  10.4  4.7
  Yes   0.0  11.3  16.0  22.0  25.4  14.4 11.0




```{r}
library(dplyr)

# Ensure election_date is a date format
ClarkCountyVoterwithAllegiantnames <- ClarkCountyVoterwithAllegiantnames %>%
  mutate(election_date = as.Date(election_date, format = "%Y-%m-%d"))

# Filter dataset to keep only the most recent record per voter_id
ClarkCountyVoterwithAllegiantnames_unique <- ClarkCountyVoterwithAllegiantnames %>%
  group_by(voter_id) %>%
  arrange(desc(election_date)) %>%  # Sort by most recent election date
  slice(1) %>%  # Keep only the first (most recent) record per voter_id
  ungroup()

# Check the new dataset
nrow(ClarkCountyVoterwithAllegiantnames_unique)  # Should be the number of unique voters




colnames(ClarkCountyVoterwithAllegiantnames_unique)
crosstabparty <- CrossTable(ClarkCountyVoterwithAllegiantnames_unique$allegiant_voter, 
                            ClarkCountyVoterwithAllegiantnames_unique$party, 
                            prop.chisq = FALSE, prop.t = FALSE, prop.r = TRUE)



# Create a contingency table of allegiant_voter vs party
crosstabparty <- table(ClarkCountyVoterwithAllegiantnames_unique$allegiant_voter, 
                       ClarkCountyVoterwithAllegiantnames_unique$party)

# Compute row-wise percentages
crosstabparty_percent_row <- prop.table(crosstabparty, margin = 1) * 100

# Print the table rounded to 1 decimal place for readability
print(round(crosstabparty_percent_row, 1))




```

partisanship percentages
      Democrat Independent American Party Libertarian Party
  No      31.5                        4.4               0.7
  Yes     38.2                        4.7               0.0
     
      Nevada Forward Party Nevada Green Party Nevada Transhumanist Party
  No                   0.0                0.1                        0.0
  Yes                  0.0                0.0                        0.0
     
      No Labels Nevada Non-Partisan Other (All Others) Republican
  No               0.0         36.3                1.8       25.2
  Yes              0.0         15.7                0.8       40.6

Of stadium voters who gave their age
 age_group   n percent
     18-29   3    0.8%
     30-44  73   19.1%
     45-59 117   30.6%
       60+ 189   49.5%
     Total 382  100.0%



 do some basic comparisons by age, vote history (especially first time voters), and party reg to compare stadium voters, in-person non-stadium voters, voters by mail, and non-voters. 

```{r first time voters}

library(dplyr)
View(ClarkCountyVoterwithAllegiantnames)
colnames(ClarkCountyVoterwithAllegiantnames)
glimpse(ClarkCountyVoterwithAllegiantnames$election_date)

# Ensure election_date is in Date format
ClarkCountyVoterwithAllegiantnames <- ClarkCountyVoterwithAllegiantnames %>%
  mutate(election_date = as.Date(election_date, format = "%Y-%m-%d"))

# Identify first-time voters by finding their earliest recorded vote
first_votes <- ClarkCountyVoterwithAllegiantnames %>%
  group_by(voter_id) %>%
  summarize(first_vote_date = min(election_date, na.rm = TRUE), .groups = "drop")

# Mark whether their first recorded vote was in November 2024
first_votes <- first_votes %>%
  mutate(first_time_voter_nov2024 = ifelse(format(first_vote_date, "%Y-%m") == "2024-11", 1, 0))

# Merge back with original data to retain voter details
ClarkCountyVoterwithAllegiantnames <- ClarkCountyVoterwithAllegiantnames %>%
  left_join(first_votes, by = "voter_id")

# Check how many first-time voters exist in November 2024
table(ClarkCountyVoterwithAllegiantnames$first_time_voter_nov2024)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
