---
title: "Youth Voters"
author: "Isabelle Jensen"
date: "2025-02-02"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
library(readr)


MD_County_Registration <- read_excel("MD_County_Registration_and_Demos includes demonimator.xlsx")
View(MD_County_Registration)


mva_youth <- read_csv("mva_youth_and_MD_county_registration (1).csv")
View(mva_youth)

data2 <- read_csv("data2.csv")
View(data2)

table(data2$county)
table(mva_youth$county)

data2_unique <- data2 %>%distinct(county, .keep_all = TRUE)
mva_youth_unique <- mva_youth %>% distinct(county, .keep_all = TRUE)


merged_data <- left_join(data2_unique, mva_youth_unique, by = "county")
final_data <- left_join(merged_data, MD_County_Registration, by = c("county" = "labels"))
View(final_data)

head(final_data)


final_data <- final_data[, colSums(!is.na(final_data)) > 0]


```


While you wait, can you take a quick look at the attached report and then dig into the attached data. For the report, I want to build on the regressions discussed toward the end of the report. I essentially want the following:
1) a bivariate regression of license attainment and 16- and 17-year- old registration rates
2) same for 18-year-old registration rates by county
3) Model with license attainment, % white, median HH income, % bachelors degree, % speak language other than English for 16- and 17-year- old registration rates
4) same as 3 but for 18-year-old registration rates

Note that some of the license data has weird gaps, just use the most recent available info (e.g. if 16-year-olds' data is only available for 2021 just use that to get the rate (we can try other approaches later). Also, note that the relevant denominator is weird too, there is a note in the MD_County_Registration_... file on how to use this. Isaiah put together the datav2 file that has the rate out of 100 (so 3.6 means 3.6%). I suggest using this as a guide but constructing everything fresh and making sure the values are in the same units (so 10% is always .1 or 10 consistently, I prefer the decimal places)





```{r making license percentages }
View(final_data)
colnames(final_data)


final_data <- final_data %>%
  mutate(sixteenlicensepercent = ((age16) / est_16_or_17_or_18_yo_2023.x)*100)
summary(final_data$sixteenlicensepercent)

final_data <- final_data %>%
  mutate(seventeenlicensepercent = ((age17) / est_16_or_17_or_18_yo_2023.x)*100)
summary(final_data$seventeenlicensepercent)

final_data <- final_data %>%
  mutate(sixandseventeenlicensepercent = (((age16+age17) / (2*est_16_or_17_or_18_yo_2023.x))*100))
summary(final_data$sixandseventeenlicensepercent)

final_data <- final_data %>%
  mutate(eighteenlicensepercent = ((age18) / est_16_or_17_or_18_yo_2023.x)*100)
summary(final_data$eighteenlicensepercent)





final_data <- final_data %>%
  mutate(allyouthlicensepercent = (((age16 + age17 + age18) / (est_16_or_17_or_18_yo_2023.x * 3)) * 100))



summary(final_data$allyouthlicensepercent)





statewide_youth_license_percent <- sum(final_data$age16 + final_data$age17 + final_data$age18, na.rm = TRUE) / 
                                  (sum(final_data$est_16_or_17_or_18_yo_2023.x, na.rm = TRUE) * 3) * 100

print(statewide_youth_license_percent)


# Calculate total licensed for each age group
total_age16 <- sum(final_data$age16, na.rm = TRUE)
total_age17 <- sum(final_data$age17, na.rm = TRUE)
total_age18 <- sum(final_data$age18, na.rm = TRUE)

# Calculate total estimated population for one age group across all counties
total_est_population <- sum(final_data$est_16_or_17_or_18_yo_2023.x, na.rm = TRUE)

# Compute statewide percentages
statewide_sixteen_license_percent <- (total_age16 / total_est_population) * 100
statewide_seventeen_license_percent <- (total_age17 / total_est_population) * 100
statewide_eighteen_license_percent <- (total_age18 / total_est_population) * 100

# Print results
statewide_sixteen_license_percent
statewide_seventeen_license_percent
statewide_eighteen_license_percent





```
percent_16_with_license
<dbl>2.322421
percent_17_with_license
<dbl>	32.1355
percent_18_with_license
<dbl>51.80923
percent_all_youth_with_license
<dbl>86.26718
		
		
		summary(final_data$sixteenlicensepercent)
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
 0.5383  2.5317  3.3501  3.3084  4.0721  6.8627 
 summary(final_data$seventeenlicensepercent)
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  10.01   32.03   39.64   38.71   45.97   59.05 
 summary(final_data$eighteenlicensepercent)
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  21.27   51.26   59.02   57.78   65.78   77.15 
 final_data <- final_data %>%
 mutate(allyouthlicensepercent = ((age16 + age17 + age18) / est_16_or_17_or_18_yo_2023.x) * 100)
summary(final_data$allyouthlicensepercent)
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  31.83   87.10  103.47   99.80  116.88  141.10 

```{r 2021 bivariate no controls}
colnames(final_data)
View(final_data)
youth2021 <- read_csv("youthvoters2021.csv")

sixteenjustprereg<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ sixteenlicensepercent, data = final_data)
summary(sixteenjustprereg)
#estimate 0.030662
#p 0.000285
#very significant


seventeenjustprereg<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ seventeenlicensepercent, data = final_data)
summary(seventeenjustprereg)
#estimate 0.004484
#p 5.46e-07
#significant


sixandsevenjustprereg<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ sixandseventeenlicensepercent, data = final_data)
summary(sixandsevenjustprereg)
#estimate 0.008194
#p 5.14e-07
#very significant


eighteenjustreg<-lm(est_reg_rate_18_yo_as_of_202312.x ~ eighteenlicensepercent, data = final_data)
summary(eighteenjustreg)
#estimate 0.0043047
#p 0.000128
#very significant

```





```{r 2021 regressions with controls}

sixteenwithcontrols<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ sixteenlicensepercent+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = final_data)
summary(sixteenwithcontrols)
#only white is significant at p 0.0519 positive coefficient 


seventeenwithcontrols<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ seventeenlicensepercent+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = final_data)
summary(seventeenwithcontrols)
#only license percentage remains significant with p value 0.0104 and coefficient 4.490e-03

sixandseventeenwithcontrols<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ sixandseventeenlicensepercent+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = final_data)
summary(sixandseventeenwithcontrols)
#only license percentage remains significant with p value 0.00866 and coefficient 8.544e-03 


eighteenwithcontrols<-lm(est_reg_rate_18_yo_as_of_202312.x ~ eighteenlicensepercent+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = final_data)
summary(eighteenwithcontrols)
#nothing significant



```



```{r only 2023 bivariates}
str(final_data$year) 
merged_data <- left_join(data2, mva_youth, by = "county")
alljoin <- left_join(merged_data, MD_County_Registration, by = c("county" = "labels"))
colnames(alljoin)
data2023 <- alljoin %>% filter(year.x == 2023 & year.y == 2023)
View(data2023)
head(data2023)


data2023 <- data2023 %>%
  mutate(seventeenlicensepercent2023 = ((age17) / est_16_or_17_or_18_yo_2023.x)*100)
head(data2023$seventeenlicensepercent2023)


data2023 <- data2023 %>%
  mutate(eighteenlicensepercent2023 = ((age18) / est_16_or_17_or_18_yo_2023.x)*100)
head(data2023$eighteenlicensepercent2023)


seventeenjustprereg2023<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ seventeenlicensepercent2023, data = data2023)
summary(seventeenjustprereg2023)
#coefficient 0.0044902
#p value 1.91e-06
#very significant


eighteenjustreg2023<-lm(est_reg_rate_18_yo_as_of_202312.x ~ eighteenlicensepercent2023, data = data2023)
summary(eighteenjustreg2023)
#coefficient 0.0047481
#p value 2.60e-05
#very significant


seventeencontrols2023<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ seventeenlicensepercent2023+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = data2023)
summary(seventeencontrols2023)
#only license percentage remains significant with p value 0.0284  and coefficient 3.820e-03 


eighteencontrols2023<-lm(est_reg_rate_18_yo_as_of_202312.x ~ eighteenlicensepercent2023+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = data2023)
summary(eighteencontrols2023)
#only license percentage remains significant with p value 0.042018  and coefficient 4.478e-03


youth2023 <- read_csv("youthvoters2023.csv")
write.csv(data2023, "youthvoters2023.csv", row.names = FALSE)


summary(data2023$seventeenlicensepercent2023)
summary(data2023$eighteenlicensepercent2023)



state_summary <- data2023 %>%
  summarise(
    total_est_youth = sum(est_16_or_17_or_18_yo_2023.x, na.rm = TRUE),
    total_age16 = sum(age16, na.rm = TRUE),
    total_age17 = sum(age17, na.rm = TRUE),
    total_age18 = sum(age18, na.rm = TRUE)
  )

state_summary <- state_summary %>%
  mutate(
    percent_16_with_license = (total_age16 / total_est_youth) * 100,
    percent_17_with_license = (total_age17 / total_est_youth) * 100,
    percent_18_with_license = (total_age18 / total_est_youth) * 100,
    percent_all_youth_with_license = ((total_age16 + total_age17 + total_age18) / total_est_youth) * 100
  )


data2023 <- data2023 %>%
  mutate(sevenandeightteenlicensepercent = (((age17+age18) / (2*est_16_or_17_or_18_yo_2023.x))*100))

summary(data2023$sevenandeightteenlicensepercent)

state_summary <- data2023 %>%
  summarise(
    total_est_youth_17_18 = sum(est_16_or_17_or_18_yo_2023.x, na.rm = TRUE) * (2/3),  # Approximate 17-18 population
    total_17_18_with_license = sum(age17 + age18, na.rm = TRUE)  # Total 17-18 licenses
  )
state_summary <- state_summary %>%
  mutate(percent_17_18_with_license = (total_17_18_with_license / total_est_youth_17_18) * 100)

print(state_summary)


state_summary <- final_data %>%
  summarise(
    total_est_youth_17_18 = sum(est_16_or_17_or_18_yo_2023.x, na.rm = TRUE) * (2/3),  # Approximate 17-18 population
    total_17_18_with_license = sum(age17 + age18, na.rm = TRUE)  # Total licenses for 17-18 year olds
  )
state_summary <- state_summary %>%
  mutate(percent_17_18_with_license = (total_17_18_with_license / total_est_youth_17_18) * 100)
print(state_summary)


```
percent_17_with_license
<dbl>
percent_18_with_license
<dbl>
percent_all_youth_with_license
<dbl>
0	26.56452	49.27406	75.83858




2023 seventeen 
 Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  8.089  26.479  33.225  33.318  41.429  53.561 
 2023 eighteen
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  19.57   46.70   57.21   54.88   64.02   71.66 
  
  17 and 18
    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  13.83   35.82   46.31   44.10   52.08   62.61 
  
```{r graphs}

colnames(final_data)
sixandsevenjustprereg<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ sixandseventeenlicensepercent, data = final_data)
summary(sixandsevenjustprereg)

View(final_data)
glimpse(final_data$est_prereg_rate_16_17_yo_as_of_202312.x)
final_data$est_prereg_rate_16_17_yo_as_of_202312.percentage <- final_data$est_prereg_rate_16_17_yo_as_of_202312.x * 100

library(ggplot2)
library(ggrepel)
library(stringr)

# Example scatterplot with regression line and labeled points
ggplot(final_data, aes(x = sixandseventeenlicensepercent, y = est_prereg_rate_16_17_yo_as_of_202312.percentage, label = county)) +
  geom_point() +  
  geom_smooth(method = "lm", color = "blue") +  # Add regression line
  geom_text(vjust = -0.5, hjust = 0.5, size = 3)+  # Label points
  theme_minimal() +  # Use a clean theme
  labs(x = "Percent of 16 and 17 year olds with license", y = "Preregistration Rates of 16 and 17 year olds", title = "Scatterplot with Regression Line and Labels")



final_data$countynice <- str_replace_all(final_data$county, "_", " ")  # Replace underscores
final_data$countynice <- str_to_title(final_data$countynice)  # Capitalize each word

# Check the result
glimpse(final_data$countynice)

# Create the scatterplot with better label positioning
ggplot(final_data, aes(x = sixandseventeenlicensepercent, y = est_prereg_rate_16_17_yo_as_of_202312.percentage, label = countynice)) +
  geom_point() +  
  geom_smooth(method = "lm", color = "blue") +  # Add regression line
  geom_text_repel(size = 3, box.padding = 0.25, point.padding = 0.3) +  # Improve label spacing
  theme_gray() +  # Use a clean theme
  labs(x = "Percent of 16 and 17 Year Olds with License", 
       y = "Preregistration Percentage of 16 and 17 Year Olds", 
       title = "Relationship Between 16 and 17 Year Old License Holding and Preregistration")


data2023$countynice <- str_replace_all(data2023$county, "_", " ")  # Replace underscores
data2023$countynice <- str_to_title(data2023$countynice)
data2023$est_reg_rate_18_yo_as_of_202312.percentage <- data2023$est_reg_rate_18_yo_as_of_202312.x * 100
ggplot(data2023, aes(x = eighteenlicensepercent2023, y = est_reg_rate_18_yo_as_of_202312.percentage, label = countynice)) +
  geom_point() +  
  geom_smooth(method = "lm", color = "blue") +  # Add regression line
  geom_text_repel(size = 3, box.padding = 0.25, point.padding = 0.3) +  # Improve label spacing
  theme_gray()  +  # Use a clean theme
  labs(x = "Percent of 18 Year Olds with License", 
       y = "Registration Percentage of 18 Year Olds", 
       title = "Relationship Between 18 Year Old License Holding and Registration")

```





```{r, echo=FALSE, fig.show="hold"}

ggplot(final_data, aes(x = sixandseventeenlicensepercent, y = est_prereg_rate_16_17_yo_as_of_202312.percentage, label = countynice)) +
  geom_point() +  
  geom_smooth(method = "lm", color = "blue") +  # Add regression line
  geom_text_repel(size = 3, box.padding = 0.25, point.padding = 0.3) +  # Improve label spacing
  theme_gray() +  # Use a clean theme
  labs(x = "Percent of 16 and 17 Year Olds with License", 
       y = "Preregistration Percentage of 16 and 17 Year Olds", 
       title = "Relationship Between 16 and 17 Year Old License Holding and Preregistration")



data2023$est_prereg_rate_16_17_yo_as_of_202312.percentage <- data2023$est_prereg_rate_16_17_yo_as_of_202312.x * 100
ggplot(data2023, aes(x = seventeenlicensepercent2023, y = est_prereg_rate_16_17_yo_as_of_202312.percentage, label = countynice)) +
  geom_point() +  
  geom_smooth(method = "lm", color = "blue") +  # Add regression line
  geom_text_repel(size = 3, box.padding = 0.25, point.padding = 0.3) +  # Improve label spacing
  theme_gray()  +  # Use a clean theme
  labs(x = "Percent of 17 Year Olds with License", 
       y = "Registration Percentage of 17 Year Olds", 
       title = "Relationship Between 17 Year Old License Holding and Registration")


ggplot(data2023, aes(x = eighteenlicensepercent2023, y = est_reg_rate_18_yo_as_of_202312.percentage, label = countynice)) +
  geom_point() +  
  geom_smooth(method = "lm", color = "blue") +  # Add regression line
  geom_text_repel(size = 3, box.padding = 0.25, point.padding = 0.3) +  # Improve label spacing
  theme_gray()  +  # Use a clean theme
  labs(x = "Percent of 18 Year Olds with License", 
       y = "Registration Percentage of 18 Year Olds", 
       title = "Relationship Between 18 Year Old License Holding and Registration")





```