---
title: "Youth Voters"
author: "Isabelle Jensen"
date: "2025-02-02"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
library(readr)


MD_County_Registration <- read_excel("MD_County_Registration_and_Demos includes demonimator.xlsx")
View(MD_County_Registration)


mva_youth <- read_csv("mva_youth_and_MD_county_registration (1).csv")
View(mva_youth)

data2 <- read_csv("data2.csv")
View(data2)

table(data2$county)
table(mva_youth$county)

data2_unique <- data2 %>%distinct(county, .keep_all = TRUE)
mva_youth_unique <- mva_youth %>% distinct(county, .keep_all = TRUE)


merged_data <- left_join(data2_unique, mva_youth_unique, by = "county")
final_data <- left_join(merged_data, MD_County_Registration, by = c("county" = "labels"))
View(final_data)

head(final_data)


final_data <- final_data[, colSums(!is.na(final_data)) > 0]


```


While you wait, can you take a quick look at the attached report and then dig into the attached data. For the report, I want to build on the regressions discussed toward the end of the report. I essentially want the following:
1) a bivariate regression of license attainment and 16- and 17-year- old registration rates
2) same for 18-year-old registration rates by county
3) Model with license attainment, % white, median HH income, % bachelors degree, % speak language other than English for 16- and 17-year- old registration rates
4) same as 3 but for 18-year-old registration rates

Note that some of the license data has weird gaps, just use the most recent available info (e.g. if 16-year-olds' data is only available for 2021 just use that to get the rate (we can try other approaches later). Also, note that the relevant denominator is weird too, there is a note in the MD_County_Registration_... file on how to use this. Isaiah put together the datav2 file that has the rate out of 100 (so 3.6 means 3.6%). I suggest using this as a guide but constructing everything fresh and making sure the values are in the same units (so 10% is always .1 or 10 consistently, I prefer the decimal places)


```{r making correct variables}
colnames(final_data)


final_data <- final_data %>%
  mutate(sixteenlicensepercent = ((age16) / est_16_or_17_or_18_yo_2023.x)*100)
head(final_data$sixteenlicensepercent)


final_data <- final_data %>%
  mutate(seventeenlicensepercent = ((age17) / est_16_or_17_or_18_yo_2023.x)*100)
head(final_data$seventeenlicensepercent)


final_data <- final_data %>%
  mutate(sixandseventeenlicensepercent = (((age16+age17) / (2*est_16_or_17_or_18_yo_2023.x))*100))

head(final_data$sixandseventeenlicensepercent)


final_data <- final_data %>%
  mutate(eighteenlicensepercent = ((age18) / est_16_or_17_or_18_yo_2023.x)*100)
head(final_data$eighteenlicensepercent)

```

```{r making correct bivariate no controls}
colnames(final_data)




sixteenjustprereg<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ sixteenlicensepercent, data = final_data)
summary(sixteenjustprereg)
#estimate 0.030662
#p 0.000285
#very significant


seventeenjustprereg<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ seventeenlicensepercent, data = final_data)
summary(seventeenjustprereg)
#estimate 0.004484
#p 5.46e-07
#significant


sixandsevenjustprereg<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ sixandseventeenlicensepercent, data = final_data)
summary(sixandsevenjustprereg)
#estimate 0.008194
#p 5.14e-07
#very significant


eighteenjustreg<-lm(est_reg_rate_18_yo_as_of_202312.x ~ eighteenlicensepercent, data = final_data)
summary(eighteenjustreg)
#estimate 0.0043047
#p 0.000128
#very significant

```

```{r making correct bivariate with controls}

sixteenwithcontrols<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ sixteenlicensepercent+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = final_data)
summary(sixteenwithcontrols)
#only white is significant at p 0.0519 positive coefficient 


seventeenwithcontrols<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ seventeenlicensepercent+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = final_data)
summary(seventeenwithcontrols)
#only license percentage remains significant with p value 0.0104 and coefficient 4.490e-03

sixandseventeenwithcontrols<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ sixandseventeenlicensepercent+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = final_data)
summary(sixandseventeenwithcontrols)
#only license percentage remains significant with p value 0.00866 and coefficient 8.544e-03 


eighteenwithcontrols<-lm(est_reg_rate_18_yo_as_of_202312.x ~ eighteenlicensepercent+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = final_data)
summary(eighteenwithcontrols)
#nothing significant



```



```{r only 2023}
str(final_data$year) 
merged_data <- left_join(data2, mva_youth, by = "county")
alljoin <- left_join(merged_data, MD_County_Registration, by = c("county" = "labels"))
colnames(alljoin)
data2023 <- alljoin %>% filter(year.x == 2023 & year.y == 2023)
View(data2023)
head(data2023)


data2023 <- data2023 %>%
  mutate(seventeenlicensepercent2023 = ((age17) / est_16_or_17_or_18_yo_2023.x)*100)
head(data2023$seventeenlicensepercent2023)


data2023 <- data2023 %>%
  mutate(eighteenlicensepercent2023 = ((age18) / est_16_or_17_or_18_yo_2023.x)*100)
head(data2023$eighteenlicensepercent2023)


seventeenjustprereg2023<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ seventeenlicensepercent2023, data = data2023)
summary(seventeenjustprereg2023)
#coefficient 0.0044902
#p value 1.91e-06
#very significant


eighteenjustreg2023<-lm(est_reg_rate_18_yo_as_of_202312.x ~ eighteenlicensepercent2023, data = data2023)
summary(eighteenjustreg2023)
#coefficient 0.0047481
#p value 2.60e-05
#very significant


seventeencontrols2023<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ seventeenlicensepercent2023+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = data2023)
summary(seventeencontrols2023)
#only license percentage remains significant with p value 0.0284  and coefficient 3.820e-03 


eighteencontrols2023<-lm(est_reg_rate_18_yo_as_of_202312.x ~ eighteenlicensepercent2023+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = data2023)
summary(eighteencontrols2023)
#only license percentage remains significant with p value 0.042018  and coefficient 4.478e-03


```














Don't bother reading. This was all wrong kept it just in case.
```{r bivariate using wrong propotions}
final_data <- final_data %>%
  mutate(percentage16and17outoftotal = ((age16 + age17) / total_md_drivers.y)*100)
head(final_data$percentage16and17outoftotal)

final_data <- final_data %>%
  mutate(percentage18outoftotal = ((age18) / total_md_drivers.y)*100)
head(final_data$percentage18outoftotal)


final_data <- final_data %>%
  mutate(percentage16and17outofyouth = ((age16 + age17) / (age16+age17+age18))*100)
head(final_data$percentage16and17outofyouth)

final_data <- final_data %>%
  mutate(percentage18outofyouth = ((age18) / (age16+age17+age18))*100)
head(final_data$percentage18outofyouth)

colnames(final_data)

just1617outoftotal<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ percentage16and17outoftotal, data = final_data)
summary(just1617outoftotal)
#estimate 0.20222
#p value 0.00044
#very significant

just1617outofyouth<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ percentage16and17outofyouth, data = final_data)
summary(just1617outofyouth)
#estimate 0.012915
#p 6e-05
#very significant


just18outofall<-lm(est_reg_rate_18_yo_as_of_202312.x ~ percentage18outoftotal, data = final_data)
summary(just18outofall)
#estimate 0.18823
#p 0.00776
#significant

just18outofyouth<-lm(est_reg_rate_18_yo_as_of_202312.x ~ percentage18outofyouth, data = final_data)
summary(just18outofyouth)
#-0.011833 estimate 
#p value 0.00548
#significant


controls1617total<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ percentage16and17outoftotal+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = final_data)
summary(controls1617total)
#only share white reached significance
#white 0.3047 and p 0.0394

controls1617youth<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ percentage16and17outofyouth+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = final_data)
summary(controls1617youth)
#none are significant

controls18total<-lm(est_reg_rate_18_yo_as_of_202312.x ~ percentage18outoftotal+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = final_data)
summary(controls18total)
#white and median income are significant
#white coefficient 0.4637 and p 0.002509
#median income coefficient 2.632e-06 and p 0.064959

controls18youth<-lm(est_reg_rate_18_yo_as_of_202312.x ~ percentage18outofyouth+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = final_data)
summary(controls18youth)
#none are significant

```


Note 
