---
title: "Youth Voters"
author: "Isabelle Jensen"
date: "2025-02-02"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
library(readr)


MD_County_Registration <- read_excel("MD_County_Registration_and_Demos includes demonimator.xlsx")
View(MD_County_Registration)


mva_youth <- read_csv("mva_youth_and_MD_county_registration (1).csv")
View(mva_youth)

data2 <- read_csv("data2.csv")
View(data2)

table(data2$county)
table(mva_youth$county)

data2_unique <- data2 %>%distinct(county, .keep_all = TRUE)
mva_youth_unique <- mva_youth %>% distinct(county, .keep_all = TRUE)


merged_data <- left_join(data2_unique, mva_youth_unique, by = "county")
final_data <- left_join(merged_data, MD_County_Registration, by = c("county" = "labels"))
View(final_data)

head(final_data)


final_data <- final_data[, colSums(!is.na(final_data)) > 0]


```


While you wait, can you take a quick look at the attached report and then dig into the attached data. For the report, I want to build on the regressions discussed toward the end of the report. I essentially want the following:
1) a bivariate regression of license attainment and 16- and 17-year- old registration rates
2) same for 18-year-old registration rates by county
3) Model with license attainment, % white, median HH income, % bachelors degree, % speak language other than English for 16- and 17-year- old registration rates
4) same as 3 but for 18-year-old registration rates

Note that some of the license data has weird gaps, just use the most recent available info (e.g. if 16-year-olds' data is only available for 2021 just use that to get the rate (we can try other approaches later). Also, note that the relevant denominator is weird too, there is a note in the MD_County_Registration_... file on how to use this. Isaiah put together the datav2 file that has the rate out of 100 (so 3.6 means 3.6%). I suggest using this as a guide but constructing everything fresh and making sure the values are in the same units (so 10% is always .1 or 10 consistently, I prefer the decimal places)





```{r making license percentages }
View(final_data)
colnames(final_data)


final_data <- final_data %>%
  mutate(sixteenlicensepercent = ((age16) / est_16_or_17_or_18_yo_2023.x)*100)


final_data <- final_data %>%
  mutate(seventeenlicensepercent = ((age17) / est_16_or_17_or_18_yo_2023.x)*100)


final_data <- final_data %>%
  mutate(sixandseventeenlicensepercent = (((age16+age17) / (2*est_16_or_17_or_18_yo_2023.x))*100))


final_data <- final_data %>%
  mutate(eighteenlicensepercent = ((age18) / est_16_or_17_or_18_yo_2023.x)*100)



write.csv(final_data, "youthvoters2021.csv", row.names = FALSE)

```

```{r 2021 bivariate no controls}
colnames(final_data)
View(final_data)
youth2021 <- read_csv("youthvoters2021.csv")

sixteenjustprereg<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ sixteenlicensepercent, data = final_data)
summary(sixteenjustprereg)
#estimate 0.030662
#p 0.000285
#very significant


seventeenjustprereg<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ seventeenlicensepercent, data = final_data)
summary(seventeenjustprereg)
#estimate 0.004484
#p 5.46e-07
#significant


sixandsevenjustprereg<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ sixandseventeenlicensepercent, data = final_data)
summary(sixandsevenjustprereg)
#estimate 0.008194
#p 5.14e-07
#very significant


eighteenjustreg<-lm(est_reg_rate_18_yo_as_of_202312.x ~ eighteenlicensepercent, data = final_data)
summary(eighteenjustreg)
#estimate 0.0043047
#p 0.000128
#very significant

```





```{r 2021 regressions with controls}

sixteenwithcontrols<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ sixteenlicensepercent+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = final_data)
summary(sixteenwithcontrols)
#only white is significant at p 0.0519 positive coefficient 


seventeenwithcontrols<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ seventeenlicensepercent+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = final_data)
summary(seventeenwithcontrols)
#only license percentage remains significant with p value 0.0104 and coefficient 4.490e-03

sixandseventeenwithcontrols<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ sixandseventeenlicensepercent+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = final_data)
summary(sixandseventeenwithcontrols)
#only license percentage remains significant with p value 0.00866 and coefficient 8.544e-03 


eighteenwithcontrols<-lm(est_reg_rate_18_yo_as_of_202312.x ~ eighteenlicensepercent+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = final_data)
summary(eighteenwithcontrols)
#nothing significant



```



```{r only 2023 bivariates}
str(final_data$year) 
merged_data <- left_join(data2, mva_youth, by = "county")
alljoin <- left_join(merged_data, MD_County_Registration, by = c("county" = "labels"))
colnames(alljoin)
data2023 <- alljoin %>% filter(year.x == 2023 & year.y == 2023)
View(data2023)
head(data2023)


data2023 <- data2023 %>%
  mutate(seventeenlicensepercent2023 = ((age17) / est_16_or_17_or_18_yo_2023.x)*100)
head(data2023$seventeenlicensepercent2023)


data2023 <- data2023 %>%
  mutate(eighteenlicensepercent2023 = ((age18) / est_16_or_17_or_18_yo_2023.x)*100)
head(data2023$eighteenlicensepercent2023)


seventeenjustprereg2023<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ seventeenlicensepercent2023, data = data2023)
summary(seventeenjustprereg2023)
#coefficient 0.0044902
#p value 1.91e-06
#very significant


eighteenjustreg2023<-lm(est_reg_rate_18_yo_as_of_202312.x ~ eighteenlicensepercent2023, data = data2023)
summary(eighteenjustreg2023)
#coefficient 0.0047481
#p value 2.60e-05
#very significant


seventeencontrols2023<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ seventeenlicensepercent2023+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = data2023)
summary(seventeencontrols2023)
#only license percentage remains significant with p value 0.0284  and coefficient 3.820e-03 


eighteencontrols2023<-lm(est_reg_rate_18_yo_as_of_202312.x ~ eighteenlicensepercent2023+ share_white.x + est_med_hh_income_2022_dollars.x + share_bachelors_degree_plus.x + share_language_other_than_english.x , data = data2023)
summary(eighteencontrols2023)
#only license percentage remains significant with p value 0.042018  and coefficient 4.478e-03


youth2023 <- read_csv("youthvoters2023.csv")
write.csv(data2023, "youthvoters2023.csv", row.names = FALSE)

```
```{r graphs}

colnames(final_data)
sixandsevenjustprereg<-lm(est_prereg_rate_16_17_yo_as_of_202312.x ~ sixandseventeenlicensepercent, data = final_data)
summary(sixandsevenjustprereg)

View(final_data)
glimpse(final_data$est_prereg_rate_16_17_yo_as_of_202312.x)
final_data$est_prereg_rate_16_17_yo_as_of_202312.percentage <- final_data$est_prereg_rate_16_17_yo_as_of_202312.x * 100

library(ggplot2)
library(ggrepel)
library(stringr)

# Example scatterplot with regression line and labeled points
ggplot(final_data, aes(x = sixandseventeenlicensepercent, y = est_prereg_rate_16_17_yo_as_of_202312.percentage, label = county)) +
  geom_point() +  
  geom_smooth(method = "lm", color = "blue") +  # Add regression line
  geom_text(vjust = -0.5, hjust = 0.5, size = 3)+  # Label points
  theme_minimal() +  # Use a clean theme
  labs(x = "Percent of 16 and 17 year olds with license", y = "Preregistration Rates of 16 and 17 year olds", title = "Scatterplot with Regression Line and Labels")



final_data$countynice <- str_replace_all(final_data$county, "_", " ")  # Replace underscores
final_data$countynice <- str_to_title(final_data$countynice)  # Capitalize each word

# Check the result
glimpse(final_data$countynice)

# Create the scatterplot with better label positioning
ggplot(final_data, aes(x = sixandseventeenlicensepercent, y = est_prereg_rate_16_17_yo_as_of_202312.percentage, label = countynice)) +
  geom_point() +  
  geom_smooth(method = "lm", color = "blue") +  # Add regression line
  geom_text_repel(size = 3, box.padding = 0.25, point.padding = 0.3) +  # Improve label spacing
  theme_gray() +  # Use a clean theme
  labs(x = "Percent of 16 and 17 Year Olds with License", 
       y = "Preregistration Percentage of 16 and 17 Year Olds", 
       title = "Relationship Between 16 and 17 Year Old License Holding and Preregistration")


data2023$countynice <- str_replace_all(data2023$county, "_", " ")  # Replace underscores
data2023$countynice <- str_to_title(data2023$countynice)
data2023$est_reg_rate_18_yo_as_of_202312.percentage <- data2023$est_reg_rate_18_yo_as_of_202312.x * 100
ggplot(data2023, aes(x = eighteenlicensepercent2023, y = est_reg_rate_18_yo_as_of_202312.percentage, label = countynice)) +
  geom_point() +  
  geom_smooth(method = "lm", color = "blue") +  # Add regression line
  geom_text_repel(size = 3, box.padding = 0.25, point.padding = 0.3) +  # Improve label spacing
  theme_gray()  +  # Use a clean theme
  labs(x = "Percent of 18 Year Olds with License", 
       y = "Registration Percentage of 18 Year Olds", 
       title = "Relationship Between 18 Year Old License Holding and Registration")

```





```{r, echo=FALSE, fig.show="hold"}

ggplot(final_data, aes(x = sixandseventeenlicensepercent, y = est_prereg_rate_16_17_yo_as_of_202312.percentage, label = countynice)) +
  geom_point() +  
  geom_smooth(method = "lm", color = "blue") +  # Add regression line
  geom_text_repel(size = 3, box.padding = 0.25, point.padding = 0.3) +  # Improve label spacing
  theme_gray() +  # Use a clean theme
  labs(x = "Percent of 16 and 17 Year Olds with License", 
       y = "Preregistration Percentage of 16 and 17 Year Olds", 
       title = "Relationship Between 16 and 17 Year Old License Holding and Preregistration")



data2023$est_prereg_rate_16_17_yo_as_of_202312.percentage <- data2023$est_prereg_rate_16_17_yo_as_of_202312.x * 100
ggplot(data2023, aes(x = seventeenlicensepercent2023, y = est_prereg_rate_16_17_yo_as_of_202312.percentage, label = countynice)) +
  geom_point() +  
  geom_smooth(method = "lm", color = "blue") +  # Add regression line
  geom_text_repel(size = 3, box.padding = 0.25, point.padding = 0.3) +  # Improve label spacing
  theme_gray()  +  # Use a clean theme
  labs(x = "Percent of 17 Year Olds with License", 
       y = "Registration Percentage of 17 Year Olds", 
       title = "Relationship Between 17 Year Old License Holding and Registration")


ggplot(data2023, aes(x = eighteenlicensepercent2023, y = est_reg_rate_18_yo_as_of_202312.percentage, label = countynice)) +
  geom_point() +  
  geom_smooth(method = "lm", color = "blue") +  # Add regression line
  geom_text_repel(size = 3, box.padding = 0.25, point.padding = 0.3) +  # Improve label spacing
  theme_gray()  +  # Use a clean theme
  labs(x = "Percent of 18 Year Olds with License", 
       y = "Registration Percentage of 18 Year Olds", 
       title = "Relationship Between 18 Year Old License Holding and Registration")

```